On The Impact of Atoms of Confusion in JavaScript Code

### 2021-04-19 1_Introduction.tex
text improvements
revisão do Prof. Fernando Castor

JavaScript is the most widely used language for web development and the most often adopted language on GitHub~\cite{BugsJS:ICST}. At the same time, JavaScript is infamously known to be ``the worst most popular language in the world''\footnote{https://github.com/getify/You-Dont-Know-JS/blob/2nd-ed/preface.md}. It is a weakly typed language~\cite{DynamicBehaviorJS:PLDI} that provides object-oriented, functional, and imperative constructs~\cite{JS-CodeSmells:SCAM}, in addition to features such as run-time evaluation~\cite{BugsJS:ICST,ClientSideBugsJS:TSE}, and asynchronous operations. All these 
%diverse characteristics and capabilities
features combined make JavaScript source code fertile ground for errors~\cite{BugsJS:ICST,ClientSideBugsJS:TSE} and confusion~\cite{JSGoodParts:book}.

Our research aims to help practitioners identify potentially confusing snippets of code, so that
developers could replace them with clearer code.
To achieve that, we investigate {\color{red}XYZ} JavaScript idioms that
might compromize program understanding. These idioms correspond to a set of candidates for \emph{atoms of confusion} in
JavaScript---that is, small
pieces of code that hinder developers' understanding of the source code of a program
for which there are functionally equivalent alternatives that are easier
to understand~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}. \castor{I think it would be really nice to have an example here. Maybe even a figure with a left-hand and a right-hand side. What do you say? If yes, I
'd go for post increment or assignment as value. If not a figure, the text could be as follow: An example of atom of confusion identified by previous work~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} is the use \texttt{if} statements without the accompanying curly braces. The alternative, in this case, is simply to always use the curly braces.}
Among the atoms in this work, X are shared between JavaScript and \clang/\cpplang (e.g., pre/post-increment), while Y others are specific to JavaScript, e.g., the atom Assignment as Value.
Recent work~\cite{DBLP:journals/ese/MedeirosLAAKRG19,DBLP:conf/sigsoft/GopsteinIYDZYC17} has
already explored the impact of atoms of confusion on program understanding,
and confirmed they negatively impact source code understandability.
However, they were constrained to statically typed languages (e.g., C and C++).
To the best of our knowledge, we are the first to explore atoms of confusion
in the context of a dynamically typed language (JavaScript).\castor{Review the number of atoms, X, and Y.}

To understand whether an atom of confusion introduces misunderstandings, we surveyed 140 developers from the Reddit community. Their task was to predict the output of short snippets of code, with half of them containing atoms of confusion. We found an improvement of at least 15\% in the correct answers when removing the confusing code for seven atoms (out of 9 atoms). The presence of the Comma Operator atom, for instance, can increase the average time to predict the output by 65\%, while the behavior of snippets containing the Pre-Increment atom were predicted 44\% faster.  An atom's presence---or lack of---in open-source repositories may reveal cases where developers should be more or less careful when writing code. To identify such cases, we mined the 100 most popular JavaScript projects from GitHub and found a total of X atoms in the analyzed projects. To learn more about the misunderstandings in the snippets used in the survey, we interviewed 15 experienced professional developers. Participants preferred code snippets \textit{without} an atom in 70\% of the cases, therefore corroborating to the survey findings. During the MSR and interview studies, we also explored additional atom candidates that are specific to the
JavaScript language.

>>>>

JavaScript is the most widely used language for web development and the most often adopted language on GitHub~\cite{BugsJS:ICST}. At the same time, JavaScript is infamously known to be ``the worst most popular language in the world''.\footnote{\url{https://github.com/getify/You-Dont-Know-JS/blob/2nd-ed/preface.md}} It is a weakly typed language~\cite{DynamicBehaviorJS:PLDI} that provides object-oriented, functional, and imperative constructs~\cite{JS-CodeSmells:SCAM}, in addition to features such as run-time evaluation~\cite{BugsJS:ICST,ClientSideBugsJS:TSE}, and asynchronous operations. All these 
%diverse characteristics and capabilities
features combined make JavaScript source code fertile ground for errors~\cite{BugsJS:ICST,ClientSideBugsJS:TSE} and confusion~\cite{JSGoodParts:book}.

Our research aims to help practitioners identify potentially confusing snippets of code and alternative snippets that can make the code clearer.
To achieve that, we investigate {\color{red}XYZ} JavaScript idioms that
might negatively impact program understanding. These idioms correspond to a set of candidates for \emph{atoms of confusion} in
JavaScript---that is, small
pieces of code that hinder developers' understanding of the source code of a program
for which there are functionally equivalent alternatives that are easier
to understand~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}. For instance, consider
the pair of code snippets in Listing~\ref{fig:lst01} (adapted from \textsc{NervJS/taro}).
%for presentation purpose only). 
The code on the left-hand side of the figure contains two atoms of confusion discussed in previous work focusing on the C language~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}: \emph{Logic as Control Flow} and \emph{Post-increment}. On the right-hand side, we present a cleaned version of the code, that is, without atoms of confusion. In the experiments of Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}, removing the atoms \emph{Logic as Control Flow} and \emph{Post-increment} improved the accuracy of program understanding tasks in  41\% and 34\%, respectively.

Recent work~\cite{DBLP:journals/ese/MedeirosLAAKRG19,DBLP:conf/sigsoft/GopsteinIYDZYC17,Langhout:2021:ACJ} has explored the impact of atoms of confusion on program understanding, and confirmed they negatively impact source code readability.
However, these previous studies focus exclusively on statically typed languages (e.g., \clang and \cpplang). To the best of our knowledge, we are the first to explore atoms of confusion in the context of a dynamically typed language (JavaScript).\castor{Review the number of atoms, X, and Y.}  JavaScript's programming culture differs from that of languages such as C and Java, e.g., because there are no types and due to the pervading use of callbacks. Thus, we may find that idioms and constructs that are atoms of confusion in those languages are not particularly difficult to understand for JavaScript programmers and vice-versa.
\castor{The following should probably be left out. I don't think we investigate atoms outside of those that were already studied in C. Or maybe say explicitly that we are studying a group of atoms that stems from the previous study?}
Furthermore, among the atom candidates we investigate in this work, {\color{red}X} are shared between JavaScript and \clang/\cpplang (e.g., pre/post-increment), while {\color{red}Y} others are specific to JavaScript, e.g., the atom Assignment as Value.

\begin{figure*}[thb]
\noindent\begin{minipage}{.45\textwidth}  
\begin{lstlisting}[language=JavaScript]
res = properties.reduce((s, b) => {
  b && s++
  return s
}
\end{lstlisting}
\end{minipage}\hfill
\begin{minipage}{.45\textwidth}
\begin{lstlisting}[language=JavaScript]
res = properties.reduce((s, b) => {
  if(b) {
    s = s + 1
  }
  return s
}
\end{lstlisting}
\end{minipage}
\captionof{lstlisting}{Code snippet from \textsc{NervJS/taro} project. Line 2 on the left-hand side of the figure contains two
  atoms of confusion discussed in ~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}: \emph{Logic
  as Control Flow} and \emph{Post-increment}. On the right-hand side, we present an alternative implementation without atoms of confusion. We could simplify the code on the left-hand side even more, by using \texttt{filter} and \texttt{length} instead of the \texttt{reduce} recursive pattern (e.g., \texttt{res = properties.filter(p => p).length}).}\label{fig:lst01}
\end{figure*}

To understand whether an atom of confusion introduces misunderstandings, we surveyed 140 developers from the Reddit community.\footnote{\url{https://www.reddit.com/r/javascript/comments/bw36h6/academic_research_on_confusing_code_on_javascript/}}
Their task was to predict the output of short snippets of code, with half of them containing atoms of confusion. 
\castor{Improvement of 15\% in terms of what. This is too loose.} We found an improvement of at least 15\% in the correct answers when removing the confusing code for seven atoms (out of 9 atoms). The presence of the Comma Operator atom, for instance, can increase the average time to predict the output by 65\%, while the behavior of snippets containing the Pre-Increment atom were predicted 44\% faster.  The presence or absence of atoms in open-source repositories may reveal cases where developers should be more or less careful when writing code. To identify such cases, we mined the 100 most popular JavaScript projects from GitHub and found a total of {\color{red}X} atoms in the analyzed projects. To learn more about the misunderstandings in the snippets used in the survey, we interviewed 15 experienced professional developers. Participants preferred code snippets \textit{without} an atom in 70\% of the cases, therefore corroborating with the survey findings. During the MSR and interview studies, we also explored additional atom candidates that are specific to the
JavaScript language.

### 2021-04-19 2_Background.tex
inclusao da secao de discussao
revisão do Prof. Fernando Castor

\section{Background and Related Work}
\label{back}
\subsection{Program Comprehension}\castor{It would be nice to have a table with the atom candidates similar to the one that appears in Gopstein et al. 2017.}

@@

However, one of the ways in which programs become less comprehensible occurs when programmers favor complicated syntactic constructs instead of using semantically equivalent versions---when these exist---allowed by the language or framework they are working with. Although this can be, to an extent, a personal matter, since different individuals might hold different opinions on whether a given syntax is difficult to understand, it is possible to empirically investigate if there exist programming languages' constructs or idioms that are considerably harder to understand, leading readers of the code to frequently make incorrect predictions of its behavior.

@@

One main example that occurs in many programming languages is the Change of Literal Encoding. For instance, the \clang code statement \lstinline{printf("\%d", 013);} 
often leads programmers to predict the output to be \texttt{13}, even though the correct answer is \texttt{11}. This occurs because a leading \texttt{0} in a number literal indicates that the number is in base \texttt{8}, a fact that is not only unknown to less experienced programmers, but misleading even for seasoned developers. After formulating a list of 19 potential atoms, Gopstein et al. \cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} were able to identify 15 snippets that present a statistically significant difference in comparative answer correctness when each of the 15 atoms was removed. The least confusing atom showed a 14\% boost in prediction accuracy when the atom was removed, whereas the most confusing one showed a 60\% accuracy increase. Our first study follows the empirical framework discussed in~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}, for all atoms that have a counterpart in JavaScript. Gopstein et al. \cite{DBLP:conf/msr/GopsteinZFC18} also performed a qualitative analysis on the different nature of confusion amongst the atoms, noting that, for some atoms, a particular incorrect answer was prevalent, whilst there was a distribution of incorrect ones in other atoms.

@@

@@

Castor~\cite{castor2018} presented a preliminary catalog of six atom candidates for the Swift programming language. Unlike JavaScript, in Swift most of the atoms identified by Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} are avoided by construction, e.g., it does not have assignments as values, increment operators, or macros. The paper also presents a structured definition for what it means for a code pattern to be an atom of confusion and a set of general principles to help in the identification of new atoms.

@@

Santos and Gerosa~\cite{Santos:2018:ICP} analyze the impact of eleven coding practices on program readability. Some of them, such as avoiding more than three levels of code
block nesting, are atom candidates. 
% Diego: Troquei "this work" para "their work". Removi o "more" de "are more related"
Their work evaluates the coding practices solely based on personal opinions of the study subjects. Furthermore, it focuses on the Java language. Finally, most of the practices are related to code formatting than language constructs or idioms.

>>>>

\section{Background and Related Work}
\label{back}
\subsection{Program Comprehension}

% \castor{It would be nice to have a table with the atom candidates
% similar to the one that appears in Gopstein et al. 2017.}

@@

However, one of the ways in which programs become less comprehensible occurs when programmers favor complicated syntactic constructs instead of using semantically equivalent versions---when these exist---that are potentially less complex. %allowed by the language or framework they are working with. 
This can be a matter of personal taste, since different individuals might hold different opinions on whether a given syntax is difficult to understand. Nevertheless, it is possible to empirically investigate if there exist programming language constructs or idioms that are considerably harder to understand, leading readers of the code to frequently make incorrect predictions of its behavior.

@@

One main example that occurs in many programming languages is the Change of Literal Encoding. For instance, the \clang code statement \lstinline{printf("\%d", 013);} 
often leads programmers to predict the output to be \texttt{13}, even though the correct answer is \texttt{11}. This occurs because a leading \texttt{0} in a number literal indicates that the number is in base \texttt{8}, a fact that is not only unknown to less experienced programmers, but misleading even for seasoned developers. After formulating a list of 19 atom candidates, Gopstein et al. \cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} were able to identify 15 code idioms and language constructs that present a statistically significant difference in comparative answer correctness when each of the 15 atoms was removed. The least confusing atom showed a 14\% boost in prediction accuracy when the atom was removed, whereas the most confusing one showed a 60\% accuracy increase. Our first study follows the empirical framework discussed in~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}, for all atoms that have a counterpart in JavaScript. Gopstein et al. \cite{DBLP:conf/msr/GopsteinZFC18} also performed a qualitative analysis on how confusion manifested when developers inspected code exhibiting different atoms.\castor{I'd rather have two concrete examples here, instead of this.} They note that, for some atoms, a particular incorrect answer was prevalent, whilst there was a variety of incorrect answers for other atoms.

@@

The first study of Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} inspired other studies to further investigate the impact of atoms of confusion on code comprehension.
Oliveira et al.~\cite{TheEyesDoNotLie} conducted an experiment using an eye-tracker with 30 participants involving three atoms.
They found that code with atoms of confusion requires more time from participants to predict the output and more visual effort to comprehend.
Gopstein et al., in another paper~\cite{ThinkingAloud} further explored the original atoms by additionally conducting interviews and having programmers discuss between themselves the studied atoms.
The authors argue that this method is important to complement a quantitative analysis.
Indeed, their study revealed findings such as a \textit{``correct evaluation of an atom might not mean that a programmer understood its meaning''
}. \castor{We use interviews, but did we "test" the interviewees to verify if they understand code?} We follow a similar approach as we augment our quantitative analysis with interviews.

Gopstein et al. also influenced researchers to look into atoms in languages other than \clang and \cpplang.
Langhout and Aniche~\cite{Langhout:2021:ACJ} derived a set of 14 atoms for Java and performed an experiment with 132 students. They found that participants are more likely to make mistakes (ranging from 2.7 up to 56 times) when atoms are present in half of their studied atoms.
Castor~\cite{castor2018} presented a preliminary catalog of six atom candidates for the Swift programming language. Unlike JavaScript, in Swift, most of the atoms identified by Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} are avoided by construction, e.g., it does not have assignments as values, increment operators, or macros. The paper also presents a structured definition for what it means for a code pattern to be an atom of confusion and a set of general principles to help in the identification of new atoms.

@@

% Castor~\cite{castor2018} presented a preliminary catalog of six atom candidates for the Swift programming language. Unlike JavaScript, in Swift most of the atoms identified by Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} are avoided by construction, e.g., it does not have assignments as values, increment operators, or macros. The paper also presents a structured definition for what it means for a code pattern to be an atom of confusion and a set of general principles to help in the identification of new atoms.

@@

Santos and Gerosa~\cite{Santos:2018:ICP} analyze the impact of eleven coding practices on program readability. Some of them, such as avoiding more than three levels of code block nesting, are atom candidates. 
% Diego: Troquei "this work" para "their work". Removi o "more" de "are more related"
Their work evaluates the coding practices solely based on personal opinions of the study subjects. Furthermore, it focuses on the Java language. Finally, most of the practices analyzed in this study are more related to code formatting than language constructs or idioms.

### 2021-04-19 3_Method.tex
revisão do Prof. Fernando Castor

\section{Study Settings}
\label{method}

The main goal of this research is to investigate the impact of atoms of confusion in JavaScript code comprehension. As such, in this paper we answer the following research questions:

@@

Answering these research questions has several implications. For instance, we can either generalize or refute the perceptions about atoms of confusion already discussed in the literature (goal of research questions (RQ1) and (RQ2)). In addition, answers to the third and fourth questions allow us to enrich existing catalogs about atoms of confusion and discuss how often they occur in practice. Answering these research questions also lay the foundations for the implementation of tools that can automatically transform code into its cleaner versions---though we postpone these results to a future research work. To answer these research questions we conduct a mixed-methods study, including a survey, a set of interviews, and an effort of mining JavaScript software repositories.

\subsection{Survey Study Settings}

@@

The design of our first study blocks two variables
(developer experience and the code snippets) and
considers two treatments: the presence or absence of 
atoms of confusion within the code snippets. 
To achieve such a design goal, controlling the effect of experience and individual code snippets, we resorted to the \textit{Latin Square Design} \cite{Hunter-Experimenters}. Using this design we create a 2 x 2 matrix in which each row represents a subject and each column indicates the set of code snippets. The design of each square (a replica) is such that no treatment is repeated in the same row or column. For example, considering that we have a set of 10
code samples $s_0, s_1, ..., s_{10}$, if a given subject (P1) is asked to predict the output of the code samples $s_0, s_1, ..., s_5$ that contain atom candidates, then, when answering questions about non-confusing code snippets, they will only be presented with non-confusing versions of the code samples $s_6, s_7,..., s_{10}$. Furthermore, a given subject (P2), which constitutes the second row of our example square, will be asked questions about the non-confusing versions for $s_1, s_2, ..., s_5$, and will answer questions about confusing snippets of for $s_6, s_7,..., s_{10}$. By doing that, we guarantee that all versions of code snippets are contained within each square, and that each configuration occurs only once within a square. Figure \ref{fig:latinsquare} offers a visual representation of the concept.

  \begin{figure}[htb!]
      \noindent
      \centering
      \includegraphics[scale=.50]{images/latin-square.pdf}
      \caption{Latin square design. Each ``square'' corresponds to 
      a replica in our study. Each replica comprises two participants (square rows) 
      and two sets of code snippets (CS-1 and CS-2). We randomly apply the 
      treatments (atom or non-atom code) to the cells of the squares.} 
      \label{fig:latinsquare}
  \end{figure}

Having selected 10 atoms in our first study, we wrote small pieces of code that contained each atom. We also wrote their equivalent snippet without the confusing idioms and constructs, leading to a total of 20 code snippets. In order to reduce the cognitive effort, we decided that each subject would be asked to predict the output of a subset of 10 code snippets, wherein each subset contained five blocks of code that contained atoms of confusion, whilst the remaining five had the atoms removed. The order in which the questions were presented was randomized. By doing this, we were seeking to minimize the chances of subjects being aware that the current listing they were analyzing contained (or not) atoms of confusion.
That is, each respondent of the survey should indicate what would be the outcomes of the code snippets, some of them having atoms of confusion (while other code snippets did not). 
We measured answer correctness and the total time each participant needed to answer the survey's questions.

@@

We implemented our survey as a web application. As part of this 
effort, we carried out an informal pilot whose main objectives were: to spot bugs in the application and in the data collection mechanism; to gain feedback from respondents about the user experience of the application; and to formulate an estimate about how long answering the survey would take on average. Fellow undergraduate students, professional colleagues, and friends took the pilot survey. Some users reported layout defects, and many reported that the landing page did not explain the survey well enough. We also spotted minor issues with our routines to create and populate the Latin Squares.

We organized the survey in three sections. The first section aims to characterize the subjects, asking for their age, education level, and programming experience. We also included a check button, whose checking meant users agreed that all collected data would be used solely for research purposes. In the second section, we presented to the participants a small set of instructions, where we explained how the survey worked and asked them to dedicate their attention to it. We stressed to participants the importance of not using any aids during the survey, such as online or console interpreters. For each question page, we kept track of whether or not the subjects switched windows.

The last section of the survey presented a sequence of 10 questions, each containing a code snippet. For each question, there was a text box where the answer should be written. There was also an ``I do not know'' button, which, when clicked, led the subject to the next question. In our setting, ``I do not know'' was treated as a wrong answer. The code snippets were presented as images copied from a text editor, so as to demotivate respondents from resorting to external resources by copying and pasting the code into an interpreter. Upon submitting their answer for a particular question, the subject was automatically led to a similar page, containing another snippet.

@@

\subsection{Interview Study Settings}

To complement our initial survey, we performed semi-structured interviews with professional JavaScript developers, aiming to identify their perceptions regarding code snippets containing atoms of confusion (research question RQ2). We also asked each participant if they knew of any other JavaScript-specific construct that they regarded as being likely to make the code hard to understand (research question RQ3). In this section we detail the protocol we followed to conduct the interviews and to analyze the results.

@@

\subsubsection*{Participants Selection} We invited the participants of the interviews using a snowballing technique. That is, starting from our network of contacts, we invited an initial set of candidates to take part in our survey. From this initial list, we asked for an indication of additional candidates. Our main selection criterion was that all participants should have been working with JavaScript in their daily professional activities. We invited a total of 17 developers, and 15 of them agreed to participate.
%We conducted the interviews for a period of two weeks.

\subsubsection*{Interview Process} We conducted semi-structured interviews using a web conferencing software. We recorded all the interviews with the consent of the participants. On average, the interviews lasted 26.29 minutes, with the shortest one lasting 14.59 minutes and the longest one 43.06 minutes. The main goal of the interviews was to have developers assess whether two versions of code (one with a single atom and another one without the atom) with the same behavior differed concerning program comprehension, and, if so, which version they\
 regarded as the easier one to understand. Two of the researchers conducted the interviews, and a third one listened to all the recordings to cross-validate the collected data.

The interviews had three main parts. In the first one, we asked the developers the following demographic information: name, email, gender, level of education, current job position, JavaScript experience in years and other programming languages they have worked with.
Table \ref{pinterview} summarizes this demographic information.

@@

In the second part of the interview our aim was to allow the subjects to describe
their JavaScript experience, as well as to allow them to reveal any JavaScript's constructs they regarded as innately confusing. This would allow us to identify potential atom candidates that are more specific to the JavaScript language. Examples of questions we explored in this section include: \emph{Do you think JavaScript is a language that produces code that is hard to understand?} and \emph{Do you regard any particular construct or idiom of the language as
  especially confusing?}

In the third part of the interview, participants were shown the code snippets that were used in the first study. The participants were told both sides of each pair had the same output, and were asked to evaluate which version of the code was easier to understand. The interviewers restrained themselves from introducing bias into the answers by not explaining that one of the sides of each pair contained the atom under investigation. Subjects were just presented the snippets and allowed to take the necessary time to decide on the most readable snippet.

@@

We first transcribed each interview and then examined the broad distribution of the answers. Our goal was to build an initial understanding of the participants' perceptions with regards to the challenges to understand JavaScript code in general and JavaScript code with atoms of confusion. We next followed with an open-coding	procedure, highlighting	the main themes and quoting the participant answers, which we show the most important ones in
Section~\ref{sec:interview-results}. Finally, we consolidate the feedback of the participants with respect to the confusion that our set of atom candidates might introduce in JavaScript code.

@@

\subsection{Mining JavaScript Software Repositories Settings}

To understand how often atoms of confusion appear in open source projects, and thus answer our fourth research question (\emph{What is the frequency of occurrence of atoms of confusion in practice?}), we mined a set of GitHub open source repositories. To this end, we first collected the most popular GitHub repositories that are primarily written in JavaScript. We measured popularity using the project's stargazers. This metric, available on  GitHub API, represents the number of stars a project received from users of the platform. The same metric has been used in a number of previous studies as a proxy to estimate project's popularity~\cite{gyimesi2019bugsjs,canedo:esem2020}. We then selected the top 100 most popular repositories and removed projects that did not reach the first quartile of the distribution of lines of code.

@@

We mined the occurrence of atom candidates from the repositories in our curate dataset using source code queries that we had written using the CodeQL language~\cite{moor:gttse2007}. CodeQL is an object-oriented variant of the Datalog language~\cite{rodriguez2020efficient}, and currently supports researchers and practitioners to query the source code of systems written in different languages (such as \cpplang, Java, and JavaScript). We also automate the process of running the queries and exporting the results to a format that simplifies our
analysis (and also the reproduction of this study). Finally, we carried
out basic descriptive statistics, in order to understand how often the
atoms of confusion appear in practice.

>>>>

\section{Methodology}
\label{method}

The main goal of this research is to investigate the impact of atoms of confusion on JavaScript code comprehension. As such, in this paper we answer the following research questions:

@@

Answering these research questions has several implications. For instance, we can either generalize or refute the perceptions about atoms of confusion already discussed in the literature (goal of research questions (RQ1) and (RQ2)). In addition, answers to the third and fourth questions allow us to enrich existing catalogs about atoms of confusion and discuss how often they occur in practice. Answering these research questions also lay the foundations for the implementation of tools that can automatically transform code into its cleaner versions---though we postpone these results to a future research work. To answer these research questions we conduct a mixed-methods study, including a survey, a set of interviews, and a JavaScript software repository mining study.

\subsection{Survey Study Setting}\label{sec:meth:survey}

@@

The design of our first study blocks two variables
(developer experience and the code snippets) and
considers two treatments: the presence or absence of 
atoms of confusion within the code snippets. 
To achieve such a design goal, controlling the effect of experience and individual code snippets, we resorted to the \textit{Latin Square Design} \cite{Hunter-Experimenters}. Using this design we create a 2 x 2 matrix in which each row represents a subject and each column indicates the set of code snippets. The design of each square (a replica) is such that no treatment is repeated in the same row or column. For example, considering that we have a set of 10
code samples $s_0, s_1, ..., s_{10}$, if a given subject (P1) is asked to predict the output of the code samples $s_0, s_1, ..., s_5$ that contain atom candidates, then, when answering questions about non-confusing code snippets, they will only be presented with non-confusing versions of the code samples $s_6, s_7,..., s_{10}$. Furthermore, a given subject (P2), which constitutes the second row of our example square, will be asked questions about the non-confusing versions for $s_1, s_2, ..., s_5$, and will answer questions about confusing snippets for $s_6, s_7,..., s_{10}$. By doing that, we guarantee that all versions of code snippets are contained within each square, and that each configuration occurs only once within a square. Figure \ref{fig:latinsquare} offers a visual representation of the concept.

  \begin{figure}[htb!]
      \noindent
      \centering
      \includegraphics[scale=.50]{images/latin-square.pdf}
      \caption{Latin square design. Each ``square'' corresponds to 
      a replica in our study. Each replica comprises two participants (square rows, e.g., P1 and P2) 
      and two sets of code snippets (CS-1 and CS-2). We randomly apply the 
      treatments (atom or non-atom code) to the cells of the squares.} 
      \label{fig:latinsquare}
  \end{figure}

Having selected 10 atom candidates in our first study, we wrote 10 short programs, one containing each atom candidate. We call them the confusing versions of the programs. We also wrote corresponding short programs without the confusing idioms and constructs and we call them the clean versions of the programs. Overall, the survey employed 20 code snippets, 10 confusing and 10 clean. In order to reduce the cognitive effort, each subject was asked to predict the output of 10 code snippets, five confusing and five clean. The order in which the questions were presented was randomized. By doing this, we were seeking to minimize the chances of subjects being aware that the current listing they were analyzing contained (or not) atoms of confusion.
That is, each respondent of the survey should indicate what would be the outcomes of the code snippets, some of them having atoms of confusion (while other code snippets did not).  We measured answer correctness and the total time each participant needed to answer the survey's questions.

@@

We implemented our survey as a web application. As part of this effort, we carried out an informal pilot whose main objectives were (i) to spot bugs in the application and in the data collection mechanism; (ii) to gain feedback from respondents about the user experience of the application; and (iii) to formulate an estimate about how long answering the survey would take on average. Fellow undergraduate students, professional colleagues, and friends took the pilot survey. Some users reported layout defects, and many reported that the landing page did not explain the survey well enough. We also spotted minor issues with our routines to create and populate the Latin Squares.

We organized the survey in three sections. The first section aims to characterize the subjects, asking their age, education level, and programming experience. We also included a check button, whose checking meant users agreed that all collected data would be used solely for research purposes. In the second section, we presented to the participants a small set of instructions, where we explained how the survey worked and asked them to dedicate their attention to it. We stressed to participants the importance of not using any aids during the survey, such as online or console interpreters. For each question page, we kept track of whether or not the subjects switched windows.

The last section of the survey presented a sequence of ten questions, each containing a code snippet. For each question, there was a text box where the answer should be written. There was also an ``I do not know'' button, which, when clicked, led the subject to the next question. In our setting, ``I do not know'' was treated as a wrong answer. The code snippets were presented as images copied from a text editor, so as to demotivate respondents from resorting to external resources by copying and pasting the code into an interpreter. Upon submitting their answer for a particular question, the subject was automatically led to a similar page, containing another snippet.

@@

\subsection{Interview Study Setting}

To complement our initial survey, we performed semi-structured interviews with professional JavaScript developers, aiming to identify their perceptions regarding code snippets containing atoms of confusion (research question RQ2). We also asked each participant if they knew of any other JavaScript-specific construct or idiom that they regarded as being likely to make the code hard to understand (research question RQ3). In this section we detail the protocol we followed to conduct the interviews and to analyze the results.

@@

\subsubsection*{Participant Selection} We invited the participants of the interviews using a snowballing technique. That is, starting from our network of contacts, we invited an initial set of candidates to take part in our survey. From this initial list, we asked for an indication of additional candidates. Our main selection criterion was that all participants should have been working with JavaScript in their daily professional activities. We invited a total of 17 developers, and 15 of them agreed to participate.
%We conducted the interviews for a period of two weeks.

\subsubsection*{Interview Process} We conducted semi-structured interviews using a web conferencing software. We recorded all the interviews with the consent of the participants. On average, the interviews lasted 26.29 minutes, with the shortest one lasting 14.59 minutes and the longest one 43.06 minutes.

%In the interviews we presented developers with two functionally equivalent short programs, one with a single atom and another one without the atom, differed concerning program comprehension. We then asked them to tell us which program was easier to understand. 
%The main goal of the interviews was to have developers assess whether two functionally equivalent short programs, one with a single atom and another one without the atom, differed concerning program comprehension, and, if so, which version they regarded as the easier one to understand.

Two of the researchers conducted the interviews, and a third one listened to all the recordings to cross-validate the collected data.
The interviews had three main parts. In the first one, we asked the developers the following demographic information: name, email, gender, level of education, current job position, JavaScript experience in years and other programming languages they have worked with.
Table \ref{pinterview} summarizes this demographic information.

@@

In the second part of the interview our aim was to allow the subjects to describe their JavaScript experience, as well as to allow them to reveal any JavaScript constructs they regarded as innately confusing. This would allow us to identify potential atom candidates that are more specific to the JavaScript language. Examples of questions we explored in this section include: \emph{Do you think JavaScript is a language that produces code that is hard to understand?} and \emph{Do you regard any particular construct or idiom of the language as
  especially confusing?}

In the third part of the interview, participants were shown the code snippets that were used in the first study. The participants were told both sides of each pair had the same output, and were asked to evaluate which version of the code was easier to understand. The interviewers avoided introducing bias in the answers by not explaining that one of the sides of each pair contained the atom under investigation. Subjects were just presented the snippets and allowed to take the necessary time to decide on the most readable snippet.

@@

We first transcribed each interview and then examined the broad distribution of the answers. Our goal was to build an initial understanding of the participants' perceptions with regards to the challenges to understand JavaScript code in general and JavaScript code with atoms of confusion in particular. We next followed with an open-coding procedure, highlighting the main themes and quoting the answers of the participants. We present these results in Section~\ref{sec:interview-results}. Finally, we consolidate the feedback of the participants with respect to the confusion that our set of atom candidates might introduce in JavaScript code.

@@

\subsection{Mining JavaScript Software Repositories Setting}

To understand how often atoms of confusion appear in open source projects, and thus answer our fourth research question (\emph{What is the frequency of occurrence of atoms of confusion in practice?}), we mined a set of GitHub open source repositories. To this end, we first collected the most popular GitHub repositories that are primarily written in JavaScript. We measured popularity using the project's stargazers. This metric, available through the GitHub API, represents the number of stars a project received from users of the platform. The same metric has been used in a number of previous studies as a proxy to estimate project's popularity~\cite{gyimesi2019bugsjs,canedo:esem2020}. We then selected the top 100 most popular repositories and removed projects that did not reach the first quartile of the distribution of lines of code.

@@

We mined the occurrence of atom candidates from the repositories in our curated dataset using source code queries that we wrote using the CodeQL language~\cite{moor:gttse2007}. CodeQL is an object-oriented variant of the Datalog language~\cite{rodriguez2020efficient}, and currently supports researchers and practitioners to query the source code of systems written in different languages (such as \cpplang, Java, and JavaScript). We also automate the process of running the queries and exporting the results to a format that simplifies our analysis (and also the reproduction of this study). Finally, we carried out basic descriptive statistics in order to understand how often the
atoms of confusion appear in practice.
